---
title: "Working with R"
subtitle: "Data Science Lifecycle"
format:
  html:
    embed-resources: true
editor: visual
editor_options: 
  chunk_output_type: console
---

## Import

We import data using packages that provide access to datasets and analysis tools. This is one way of importing data.

```{r}
library(ggplot2)
library(dplyr)
library(gapminder)
```

## Explore

The first step when working with new unknown data is to explore it and learn about the variables.

```{r}
head(gapminder)
tail(gapminder)
glimpse(gapminder)
str(gapminder)
nrow(gapminder)
ncol(gapminder)
```

## Transform - Narrow down

We change the format of the data to prepare it for analysis.

-   Keyboard shortcut for pipe operator `|>`: Ctrl / Cmd + Shift + M
-   Keyboard shortcut for assignment operator `<-`: Alt + -

```{r}
gapminder_2007 <- gapminder |> 
  filter(year == 2007)
```

## Visualize

We explore further using visualization or prepare our analysis using visualization.

```{r}
ggplot(data = gapminder_2007,
       mapping = aes(x = continent,
                     y = lifeExp)) +
  geom_boxplot()
```

## Communicate

Finally, we communicate by rendering the file and then sharing it with others.

-   Render the document to html format and fix any errors
