---
title: "Data visualisation with ggplot2"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

## Import

```{r}
library(ggplot2)
library(ggthemes)
library(ggridges)
library(palmerpenguins)
```

## Explore

```{r}
head(penguins)
str(penguins)
```

## Visualise with ggplot2

### Functions and arguments

-   functions: ggplot(), aes(), geom_point()
-   arguments: data, mapping, color

```{r}
ggplot(data = penguins,
       mapping = aes(x = flipper_length_mm, 
                     y = body_mass_g)) +
  geom_point()
```

### Aesthetic mappings

-   options: x, y, color, shape, size, alpha

```{r}
ggplot(data = penguins,
       mapping = aes(x = flipper_length_mm, 
                     y = body_mass_g, 
                     color = species,
                     shape = species)) +
  geom_point()
```

### Settings

```{r}
ggplot(data = penguins,
       mapping = aes(x = flipper_length_mm, 
                     y = body_mass_g, 
                     color = species,
                     shape = species)) +
  geom_point(size = 5, alpha = 0.7)
```

### Color scales

```{r}

ggplot(data = penguins,
       mapping = aes(x = flipper_length_mm, 
                     y = body_mass_g, 
                     color = species,
                     shape = species)) +
  geom_point(size = 5, alpha = 0.7) +
  scale_color_colorblind() 

```

```{r}

ggplot(data = penguins,
       mapping = aes(x = flipper_length_mm, 
                     y = body_mass_g, 
                     color = species,
                     shape = species)) +
  geom_point(size = 5, alpha = 0.7) +
  scale_color_manual(values = c("red", "blue", "green")) 

```

### Facets

Keyboard shortcut for the tilde (\~) varies by keyboard layout:

-   US keyboard Windows/Mac: Shift + \` (top left of your keyboard next to the 1)
-   UK keyboard Windows/Mac: Shift + \# (bottom right of your keyboard, next to Enter)
-   CH keyboard Windows/Max: Alt/Option + -

```{r}
ggplot(data = penguins,
       mapping = aes(x = flipper_length_mm,
                     y = body_mass_g)) +
  geom_point() +
  facet_grid(species ~ island)
```

```{r}
ggplot(data = penguins,
       mapping = aes(x = flipper_length_mm,
                     y = body_mass_g)) +
  geom_point() +
  facet_wrap(~species)
```

### Themes

```{r}

ggplot(data = penguins,
       mapping = aes(x = flipper_length_mm, 
                     y = body_mass_g, 
                     color = species,
                     shape = species)) +
  geom_point(size = 5, alpha = 0.7) +
  scale_color_colorblind() +
  theme_minimal()

```

## Visualising distributions

### Categorical variables

```{r}
ggplot(data = penguins,
       mapping = aes(x = species)) +
  geom_bar()
```

```{r}
ggplot(data = penguins,
       mapping = aes(x = species,
                     fill = island)) +
  geom_bar()
```

### Numerical continuous variables

```{r}
ggplot(data = penguins,
       mapping = aes(x = body_mass_g)) +
  geom_histogram()
```

```{r}
ggplot(data = penguins,
       mapping = aes(x = body_mass_g,
                     fill = species)) +
  geom_histogram()
```

```{r}
ggplot(data = penguins,
       mapping = aes(x = body_mass_g,
                     fill = species)) +
  geom_density()
```

```{r}
ggplot(data = penguins,
       mapping = aes(x = body_mass_g,
                     y = species,
                     fill = species)) +
  geom_density_ridges()
```

## Communicate

-   Render the document to html format and fix any errors
-   In your homework assignment, we will check if your document renders without errors
